A Note on Convergence Diagnostics for Multiple Imputation Algorithms

### 2020-03-12 2.ResearchReport/1.SimulationStudy/Writeup/NoteConvergence.tex
Copy simulation Research Report from master branch to merge branches

<<echo = FALSE, eval = TRUE>>=
load("../Results/results.Rdata")
names(dat) <- c("It.", "Bias", "CI width", "Cov. rate", "R chain mean", "R chain var", "AC chain mean", "AC chain var")
@

@@

% % See "3.CreateFigure.R", I specifically chose to NOT generate the figures within this Rnw file because of journal guidelines.
\begin{figure}[h]
  \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
        \includegraphics{Figures/convergence_diag.pdf}
  }
  \caption{Convergence diagnostics over 1000 MCMC simulations.}
    \label{fig:conv}
\end{figure}

% % With SEs
% \begin{figure}[h]
%   \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
%         \includegraphics{Figures/conv_diag_SEs.pdf}
%   }
%   \caption{Convergence diagnostics over 1000 MCMC simulations.}
%     \label{fig:conv2}
% \end{figure}

@@

% % With SEs
% \begin{figure}[h]
%   \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
%         \includegraphics{Figures/sim_diag_SEs.pdf}
%   }
%   \caption{Simulation diagnostics over 1000 MCMC simulations.}
%     \label{fig:sim2}
% \end{figure}

@@

CIW is only clearly divergent in the simulation condition $T=1$, see Figure \ref{fig:sim}B. However, under-estimating the variance of $\bar{Q}$, which is the case for $T=3$, may yield spurious inferences. Only conditions where $T>3$ are therefore considered sufficient. These conditions have similar CIWs, but across iterations stability is not established (i.e., the Loess line is never completely flat within the 100 simulation conditions). %After about twenty-five iterations the average CIW across all simulation conditions and repetitions is reached \Sexpr{round(mean(dat[, "CI width"]), 3)}.

The empirical coverage rate across repetitions seems more or less stable for conditions where $T>1$, see Figure \ref{fig:sim}C. We see some over-coverage at $T=2$, but that is better than under-estimating the variance of $\bar{Q}$. On average, the coverage rate is somewhat higher than the expected nominal coverage of 95\% \citep{neym34}, namely \Sexpr{100*round(mean(dat[, "Cov. rate"]), 3)}\%. Similar to the CIWs, there is some trending across iterations (i.e., the Loess line is never flat).

@@

The results in this paper were obtained using \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} \cite{R} with the \pkg{mice}~\Sexpr{paste(packageVersion("mice"), collapse = ".")} package \cite{mice}. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.

>>>>

@@

\begin{figure}[h]
  \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
        \includegraphics{Figures/convergence_diag.pdf}
  }
  \caption{Convergence diagnostics over 1000 MCMC simulations.}
    \label{fig:conv}
\end{figure}

% \begin{figure}[h]
%   \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
%         \includegraphics{Figures/conv_diag_SEs.pdf}
%   }
%   \caption{Convergence diagnostics over 1000 MCMC simulations.}
%     \label{fig:conv2}
% \end{figure}

@@

% \begin{figure}[h]
%   \resizebox{\textwidth}{!}{ %notice the \resizebox{} command
%         \includegraphics{Figures/sim_diag_SEs.pdf}
%   }
%   \caption{Simulation diagnostics over 1000 MCMC simulations.}
%     \label{fig:sim2}
% \end{figure}

@@

CIW is only clearly divergent in the simulation condition $T=1$, see Figure \ref{fig:sim}B. However, under-estimating the variance of $\bar{Q}$, which is the case for $T=3$, may yield spurious inferences. Only conditions where $T>3$ are therefore considered sufficient. These conditions have similar CIWs, but across iterations stability is not established (i.e., the Loess line is never completely flat within the 100 simulation conditions). %After about twenty-five iterations the average CIW across all simulation conditions and repetitions is reached 0.927.

The empirical coverage rate across repetitions seems more or less stable for conditions where $T>1$, see Figure \ref{fig:sim}C. We see some over-coverage at $T=2$, but that is better than under-estimating the variance of $\bar{Q}$. On average, the coverage rate is somewhat higher than the expected nominal coverage of 95\% \citep{neym34}, namely 95\%. Similar to the CIWs, there is some trending across iterations (i.e., the Loess line is never flat).

@@

The results in this paper were obtained using \proglang{R}~3.6.2 \cite{R} with the \pkg{mice}~3.7.0 package \cite{mice}. \proglang{R} itself and all packages used are available from the Comprehensive \proglang{R} Archive Network (CRAN) at \url{https://CRAN.R-project.org/}.
